A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-02 tex/codatadp.tex
more

%%% For final version as well?
\usepackage[
   a4paper,
   pdftex,
   pdftitle={A Decision Procedure for (Co)datatypes in SMT Solvers},
   pdfauthor={Andrew J. Reynolds and Jasmin Christian Blanchette},
   pdfkeywords={},
   pdfborder={0 0 0},
   draft=false,
   bookmarksnumbered,
   bookmarks,
   bookmarksdepth=2,
   bookmarksopenlevel=2,
   bookmarksopen]{hyperref}

@@

\author {Andrew J. Reynolds\inst{1} \and Jasmin Christian Blanchette\inst{2,3}}
\authorrunning {A. J. Reynolds \and J. C. Blanchette}
\institute{
\'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Switzerland
\and
Inria Nancy \& LORIA, Villers-l\`es-Nancy, France
\and
Max-Planck-Institut f\"ur Informatik, Saarbr\"ucken, Germany
}

@@

* types are simply atomic sorts, with no structure, and interpreted by nonempty
  domains

* although nothing prevents composing the decision procedure with theories
  providing polymorphic types (parametric sorts), such as the theory of arrays

@@

* in keeping with SMT conventions, predicate symbols are simply functions that
  return $\ty{Bool}$

@@

A signature consists of:
  * set of plain types, which can be interpreted or uninterpreted
    * must contain the distinguished \ty{bool} type of Booleans
  * set of function symbols with argument and return types
  * list of codatatype specifications, described below

>>>>

%%% For final version as well?
\usepackage[
   a4paper,
   pdftex,
   pdftitle={A Decision Procedure for (Co)datatypes in SMT Solvers},
   pdfauthor={Andrew Reynolds and Jasmin Christian Blanchette},
   pdfkeywords={},
   pdfborder={0 0 0},
   draft=false,
   bookmarksnumbered,
   bookmarks,
   bookmarksdepth=2,
   bookmarksopenlevel=2,
   bookmarksopen]{hyperref}

@@

\author {Andrew Reynolds\inst{1} \and Jasmin Christian Blanchette\inst{2,3}}
\authorrunning {A. Reynolds \and J. C. Blanchette}
\institute{
\'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Switzerland
\and
Inria Nancy \& LORIA, Villers-l\`es-Nancy, France
\and
Max-Planck-Institut f\"ur Informatik, Saarbr\"ucken, Germany
}

@@

* monomorphic (or many-sorted) first-order logic
* Signature: $\Sigma = (\mathcal{T}, \mathcal{F})$

* types are simply atomic sorts, with no structure, and interpreted by nonempty
  domains
  * set of types must contain the distinguished \ty{bool} type of Booleans, interpreted
    as the set of truth values, and may contain other interpreted types (e.g., \ty{int}, \ty{real}).

* although nothing prevents composing the decision procedure with theories
  providing polymorphic types (parametric sorts), such as for arrays (e.g., \ty{array}(\alpha,\beta))

* in keeping with SMT conventions, there are no separate predicate symbols;
  predicates are modeled as functions that return $\ty{Bool}$

@@

@@

We assume a signature $\Sigma = (\mathcal{T}, \mathcal{F})$.
Furthermore, a partition of the types:
  $\mathcal{T} = \mathcal{I} \uplus \mathcal{C} \uplus \mathcal{N}$
with
  * $\mathcal{I}$: datatypes
  * $\mathcal{C}$: codatatypes
  * $\mathcal{N}$: non-(co)datatypes

In practice, e.g. in an SMT problem, we'd have:
  * declaration of uninterpreted types (in any order)
  * specification of mutual groups of (co)datatypes, in a correct order
